---
title: "MA678 Final Project"
author: "Tong Sun"
date: "2022-12-11"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gridExtra)
library(readr)
library(dplyr)
library(lme4)
library(ggplot2)
library(knitr)
library(tidyverse)
library(kableExtra)
library(lmerTest)
library(car)
library(PerformanceAnalytics)
library(corrplot)
library(ggpubr)
library(psych)
library(corrgram)
library(performance)
```


# Abstract
This report focuses on the relationship between housing factors and prices, examining which factors have a large impact on prices and how they affect. In the report, I use multilevel model with different level of floors and view to examine the correlation. The square footage of the apartments interior living space has the greatest impact on the price of the house.

# Introduction
The data set I choose contains house sale prices for King County, which includes Seattle. It includes homes sold between May 2014 and May 2015. From <https://www.kaggle.com/datasets/harlfoxem/housesalesprediction>

```{r, include=FALSE}
## Data
house <- read_csv("kc_house_data.csv")
data <- house[, -c(1,2)]
```


The explanation of column names is listed below:

| column names      | explanation                                                                      |
| :--               | :-----                                                                           |
| id                | Unique ID for each home sold                                                     |
| date              | Date of the home sale                                                            |
| price             | Price of each home sold                                                          |
| bedrooms          | Number of bedrooms                                                               |
| bathrooms         | Number of bathrooms, where .5 accounts for a room with a toilet but no shower    |
| sqft_living       | Square footage of the apartments interior living space                           |
| sqft_lot          | Square footage of the land space                                                 |
| floors            | Number of floors                                                                 |
| waterfront        | A dummy variable for whether the apartment was overlooking the waterfront or not |
| view              | An index from 0 to 4 of how good the view of the property was                    |
| condition         | An index from 1 to 5 on the condition of the apartment                           |
| grade             | An index from 1 to 13, where 1-3 falls short of building construction and design, 7 has an average level of construction and design, and 11-13 have a high quality level of construction and design  |
| sqft_above        | The square footage of the interior housing space that is above ground level      |
| sqft_basement     | The square footage of the interior housing space that is below ground level      |
| yr_built          | The year the house was initially built                                           |
| yr_renovated      | The year of the house's last renovation                                          |
| zipcode           | What zipcode area the house is in                                                |
| lat               | Latitude                                                                         |
| long              | Longitude                                                                        |
| sqft_living15     | The square footage of interior housing living space for the nearest 15 neighbors |
| sqft_lot15        | The square footage of the land lots of the nearest 15 neighbors                  |


# Method

## Correlation

```{r,echo = FALSE}
corr1 <- corrplot(cor(data),        
         method = "shade", 
         type = "full",    
         diag = TRUE,      
         tl.col = "black", 
         bg = "white",    
         title = "",     
         col = NULL)   

par(mfrow=c(1,2))
corr2 <- corrplot(cor(data), method = "number",
         title = "method = 'number'",
         tl.pos = "n", mar = c(2, 1, 3, 1)) 

corr3 <- corrplot(cor(data), method = "pie",
         title = "method = 'pie'",
         tl.pos = "n", mar = c(2, 1, 3, 1)) 


``` 

First I plot the correlation plot between the 19 variable (There are total 21 variable and I remove the id and date). From the plot, some factor have high correlation with price such as sqft_living, grade and sqft_above. Some have little to do with price like condition, yr_built and latitude.

Then I choose to plot the relation between living space and grade with price at different floors and view level to see their change. From the plots, we can see that as the living space and grade of the house increases, the price of increases as well, which is the same correlation as in the previous analysis.

```{r,echo = FALSE}
living_space1 <- ggplot(data = house) + 
  aes(x = log(sqft_living), y = log(price)) + 
  geom_point(aes(color = factor(floors)), size = .5) + 
  geom_smooth(aes(color = factor(floors)), method = "lm", se = FALSE, formula = 'y ~ x')

living_space2 <- ggplot(data = house) + 
  aes(x = log(sqft_living), y = log(price)) + 
  geom_point(aes(color = factor(view)), size = .5) + 
  geom_smooth(aes(color = factor(view)), method = "lm", se = FALSE, formula = 'y ~ x')

grid.arrange(living_space1, living_space2, ncol = 2)


grade1 <- ggplot(data = house) + 
  aes(x = grade, y = log(price)) + 
  geom_point(aes(color = factor(floors)), size = .5) + 
  geom_smooth(aes(color = factor(floors)), method = "lm", se = FALSE, formula = 'y ~ x')

grade2 <- ggplot(data = house) + 
  aes(x = grade, y = log(price)) + 
  geom_point(aes(color = factor(view)), size = .5) + 
  geom_smooth(aes(color = factor(view)), method = "lm", se = FALSE, formula = 'y ~ x')

grid.arrange(grade1, grade2, ncol = 2)

```



## Model Fitting

Then we can choose some factors fit our model to analyze.

```{r}
model <- lmer(log(price) ~ bedrooms + bathrooms + log(sqft_living) + waterfront + grade + log(sqft_above) + log(sqft_living15) + (1 | floors) + (1 | view), data = data)
```


```{r, include = FALSE}
summary(model)
vif(model)
```




# Result 

## Random effects

| Groups   | Name        | Variance | Std.Dev.  |
| :--      | :--         | :--      | :--       |
| floors   | (Intercept) | 0.01460  | 0.1208    |
| view     | (Intercept) | 0.01317  | 0.1148    |
| Residual |             | 0.11070  | 0.3327    |

## Fixed effects

|                     | Estimate    | Std. Error  |df          | t value  | Pr(>|t|)     |
| :--                 | :--         | :--         | :--        | :--      | :--          |
| (Intercept)         | 8.274e+00   | 1.074e-01   | 3.997e+01  | 77.037   | < 2e-16 ***  |
| bedrooms            | -2.167e-02  | 3.238e-03   | 2.160e+04  | -6.691   | 2.27e-11 *** |
| bathrooms           | 1.681e-02   | 5.098e-03   | 2.157e+04  | 3.297    | 0.000979 *** |
| log(sqft_living)    | 4.926e-01   | 1.455e-02   | 2.160e+04  | 33.845   | < 2e-16 ***  |
| waterfront          | 3.649e-01   | 3.234e-02   | 1.367e+04  | 11.283   | < 2e-16 ***  |
| grade               | 1.947e-01   | 3.334e-03   | 2.160e+04  | 58.397   | < 2e-16 ***  |
| log(sqft_above)     | -2.313e-01  | 1.313e-02   | 2.158e+04  | -17.617  | < 2e-16 ***  |
| log(sqft_living15)  | 2.086e-01   | 1.131e-02   | 2.160e+04  | 18.451   | < 2e-16 ***  |

## Correlation of Fixed Effects
|              | (Intr)  | bedrms  | bthrms  | lg(sqft_l) | wtrfrn  | grade  | lg(sqft_b) |
| :--          | :--     | :--     | :--     | :--        | :--     | :--    | :--        |
| bedrooms     | 0.203   |         |         |            |         |        |            |
| bathrooms    | 0.217   | -0.165  |         |            |         |        |            |
| lg(sqft_lv)  | -0.214  | -0.325  | -0.384  |            |         |        |            |
| waterfront   | -0.028  | 0.042   | -0.006  | -0.004     |         |        |            |
| grade        | 0.279   | 0.167   | -0.157  | -0.177     | 0.022   |        |            |
| lg(sqft_bv)  | -0.148  | -0.025  | 0.111   | -0.577     | -0.018  | -0.165 |            |
| lg(sqft_15)  | -0.360  | 0.041   | 0.029   | -0.220     | 0.015   | -0.250 | -0.171     |


## VIF
VIF equal to 1 = variables are not correlated
VIF between 1 and 5 = variables are moderately correlated 
VIF greater than 5 = variables are highly correlated

| bedrooms  | bathrooms | log(sqft_living) | waterfront | grade     | log(sqft_above) | log(sqft_living15) |
| :--       | :--       | :--              | :--        | :--       | :--             | :--                |
| 1.643611  | 1.994794  | 5.469905         | 1.002886   | 2.059336  | 3.590107        | 2.057776           |



## Formula with fixex Effects

$$
log(price) = 8.274 - 0.02167 * bedrooms + 0.01681 * bathrooms + 0.4926 * log(sqft_living)\\ 
+ 0.3649 * waterfront + 0.1947 * grade - 0.2313 * log(sqft_above) + 0.2086 * log(sqft_living15)
$$

From the formula, we can see bedrooms and the interior housing space that is above ground level have negative effect on price and others have positive effect. Among the variables, living space are highly correlated with price.



# Discussion
From the analyze, the greatest impact on the price of house is the living space of the house, followed by the number of rooms and the level of construction and design also have a relative impact. But like the condition of the apartment and the year the house was initially built are nothing to influence.


# Appendix
```{r, echo = FALSE, warning=FALSE}
land_space1 <- ggplot(data = house) + 
  aes(x = log(sqft_lot), y = log(price)) + 
  geom_point(aes(color = factor(floors)), size = .5) + 
  geom_smooth(aes(color = factor(floors)), method = "lm", se = FALSE, formula = 'y ~ x')

land_space2 <- ggplot(data = house) + 
  aes(x = log(sqft_lot), y = log(price)) + 
  geom_point(aes(color = factor(view)), size = .5) + 
  geom_smooth(aes(color = factor(view)), method = "lm", se = FALSE, formula = 'y ~ x')

grid.arrange(land_space1, land_space2, ncol = 2)


par(mfrow=c(1,2))
res <- resid(model)
qqnorm(res)
plot(fitted(model), res)
```



